---
title: "Lessons Learned for Evidence-Based Marine Conservation: A 30-Year Perspective
  of the Mesoamerican Reef Ecoregion"
author: "Laura Givens"
date: "2025-08-14"
output:
  word_document: default
  html_document: default
subtitle: Author and topic analysis
---
```{r,message=FALSE}
library(readxl)
library(readr)
library(stringr)
library(ggplot2)
library(ggridges)
library(tidyverse)
library(patchwork)
library(flextable)
```

# 1. Download MAR Literature Review - Spreadsheet (final version: 10 Sep 2025)  

# 2. Read into R  
There is one line per paper; each line is made up of a list of all coauthor institutions 
```{r load,message=FALSE}
spdsht <- read_excel("/Users/lauragivens/Downloads/MAR\ Literature\ Review\ -\ Spreadsheet_Final.xlsx",sheet = "ReviewSummary",col_names = TRUE,skip=2,col_types = "guess")

spdsht$year <- as.double(spdsht$year)

marmp <- read_xlsx("/Users/lauragivens/Downloads/MAR\ Management\ Plan\ Review\ -\ Spreadsheet.xlsx",sheet = "Latest only",col_names = TRUE,skip=3,col_types="guess")

marmp$Year_Established <- as.double(marmp$Year_Established)
marmp$Year_ManagementPlan <- as.double(marmp$Year_ManagementPlan)
```

# 3. Subset spreadsheet to only papers that were identified as relevant  
```{r subset}
rel.spdsht <- spdsht %>% filter(str_detect(`Relevant to the manuscript?`,"Yes"))
```

# 4. Extract author countries from given addresses and split into first/coauthors  
_4a. Name repair_  
We remove the spaces in any of the country names so they aren't split later  
```{r remspaces}
rel.spdsht$addresses <- str_replace_all(rel.spdsht$addresses, "USA|US|U.S.A.|United States|United States of America", " USA")  
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "Britain|Great Britain", "Great_Britain") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "United Kingdom|UK|Englad", "UK") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "New Caledonia", "New_Caledonia") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "Hong Kong", "Hong_Kong") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "Costa Rica", "Costa_Rica") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "South Africa", "South_Africa") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "United Arab Emirates", "UAE") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "South Korea", "South_Korea") 
rel.spdsht$addresses <-   str_replace_all(rel.spdsht$addresses, "Puerto Rico", "Puerto_Rico")
rel.spdsht$addresses <- str_replace_all(rel.spdsht$addresses,"Cayman Islands", "Cayman_Islands")
rel.spdsht$addresses <- str_replace_all(rel.spdsht$addresses,"Trinidad and Tobago", "Trinidad_and_Tobago")
rel.spdsht$addresses <- str_replace_all(rel.spdsht$addresses,"Virgin Islands (British)|Virgin Islands", "Virgin_Islands")
rel.spdsht$addresses <- str_replace_all(rel.spdsht$addresses,"St Kitts & Nevi", "StKitts_and_Nevi")
```
_4b. Split strings_  
```{r splitaddresses,warning=FALSE}
author.countries.df <- rel.spdsht %>% 
  mutate(edit.addresses=addresses,.after=addresses) %>% 
  mutate(edit.addresses=sub("\\Florida.*"," USA",edit.addresses)) %>% 
  mutate(edit.addresses=sub("\\Texas.*"," USA",edit.addresses)) %>% 
  mutate(edit.addresses=sub("\\FL.*"," USA",edit.addresses)) %>% 
  mutate(edit.addresses=sub("\\TX.*"," USA",edit.addresses)) %>% 
  mutate(edit.addresses=sub("\\New York.*"," USA",edit.addresses)) %>% 
  mutate(edit.addresses=sub("\\ILLINOIS.*"," USA",edit.addresses)) %>% 
  mutate(countries_all = (strsplit(edit.addresses,"\\[[^\\]]*\\](*SKIP)(*F)|\\h*;\\h*", perl=T) %>% 
                            lapply(.,function(x) {word(x,start=-1,sep=" ")}) %>% 
                            map_chr(.,toString) 
                          ),
                          .after=edit.addresses) %>% 
  separate(countries_all,into=c("first_authors","coauthors"),sep=",",extra="merge",remove=F)
```

# 5. Data frames      

_5a. Rename some values for ease of understanding_     
_Open access tiers_  
Diamond/Platinum = OA for readers at no cost to authors   
Gold = final version of article is open access immediately following publication   
Hybrid = available to subscribers to the journal but also freely available to everyone   
Green = pre-print/publisher's version can be archived; 'self archiving'   
```{r authorcountries_updateOA}
author.countries.df <- author.countries.df %>%
  mutate(OA=case_when(open_access=="NA" ~ "Not Open Access",
                      str_detect(open_access,"iamond|latinum") ~ "Diamond",
                           str_detect(open_access,"old") ~ "Gold",
                           str_detect(open_access,"reen") & !str_detect(open_access,"old") ~ "Green",
                           str_detect(open_access,"ronze") & !str_detect(open_access,"old") & !str_detect(open_access,"reen") ~ "Bronze",
                           str_detect(open_access,"ybrid") ~ "Hybrid",
                               .default=open_access),
         .after=open_access)
```
_Shorten paper category names_
```{r authorcountries_papercategory}
author.countries.df <- author.countries.df %>% 
  mutate(PaperType=case_when(`Type of Paper` == "Basic/descriptive science" ~ "A", 
                         `Type of Paper` == "Perspective/opinion" ~ "B", 
                         `Type of Paper` == "Applied science" ~ "C", 
                         .default="Not assigned"),
         .after=`Type of Paper`
         )
```
_Identify collaboration within MAR_  
Add information about whether there was collaboration with people in country  
```{r authorcountries_marcollab}
author.countries.df <- author.countries.df %>% 
  mutate(marfirst=case_when(str_detect(first_authors,
                                       "Mexico|Belize|Guatemala|Honduras") ~ "Lead",
                            str_detect(coauthors,
                                       "Mexico|Belize|Guatemala|Honduras") ~ "Coauthor",
                            .default="Not included"),
         .after="first_authors") 
```
_Bin everything before 1990_  
Add information about whether there was collaboration with people in country  
```{r authorcountries_binyear}
author.countries.df <- author.countries.df %>% ungroup() %>% 
  mutate(Oldyear=year,
         year=case_when(year<1990 ~ 1990,.default=year)) 
```

_5b. Make figure data frames_    
Calculate how productive each country is and how that number changes over time  
```{r firstauthcountriesdf}
firstauthcountries.df <- author.countries.df %>% 
  add_count(first_authors,sort=TRUE,name="TotalPapersCountry") %>%  #count how many papers for each country
  mutate(first_authors=factor(first_authors, #set names of country as factors
                              levels=unique(first_authors),ordered=TRUE)) %>% 
  ungroup() %>% 
  group_by(year) %>% #group by year
  mutate(papers_peryear=n()) %>% #count number of total papers in a year
  ungroup() %>% group_by(year,first_authors,papers_peryear) %>% #group by year and country
  reframe(author_peryear=n(), #count number of papers coming out of a country in each year
          TotalPapersCountry=TotalPapersCountry) 
```
How many countries overall are involved in the papers we identified?  
```{r allcountriesdf}
all.countries <- lapply((author.countries.df %>% 
  separate_wider_delim(., #split a string into columns 
                       cols=countries_all,
                       delim=",", #delimiter between values to separate
                       names_sep="_", #output column names are separated with
                       too_few="align_start" #adds NA to any strings that are too short to fill in all columns
                       ) %>% 
    select(starts_with("countries"))),unique) %>% 
  c() %>% unlist() %>% 
  trimws(.,which="left") %>% unique()
```
data frame to look at the country affiliation of ALL authors (first and coauthor)  
```{r yearcountriesdf}
 year.countries.df <- author.countries.df %>% 
  separate_wider_delim(.,cols=countries_all, #separate the list of all author countries 
                       delim=",", #separate by comma
                       names_sep="_", #delimiter separating names of new columns
                       too_few="align_start") %>% #if there are uneven numbers, fill the remaining with NA  
  select(c(year,Oldyear,starts_with("Type"),language,starts_with("countr"))) %>% #keep only year, type of paper, language, and country columns  
  #group_by(year) %>% summarize(across(starts_with("count"),n_distinct))
pivot_longer(cols=c(starts_with("countries")), #take all country columns and make each their own row. makes a new column 'name' with the column name each country value came from 
              #names_to="year",
              values_drop_na=TRUE
              ) %>% 
  mutate(value=trimws(value,which="left")) %>% 

    #first, sum the country affiliation of each first author for each year 

  group_by(year,value,name) %>% 
  mutate(firstauth=case_when(name=="countries_all_1" & row_number() == 1 ~ n(), #get the count of number of first author papers for each country each year
                             .default=NA)) %>%
  
  #next, sum the total participation: for ALL authors, how many are from each country each year?
  
  ungroup() %>%
  group_by(year,value) %>% 
  mutate(total_percountry_peryear = case_when(
    row_number() == 1 ~ n(),
    .default=NA
  )) %>% #distinct() %>%
  #group_by(year,value) %>%  #filter(n()>1 & !is.na(firstauth) | n()==1) %>% #filter out duplicate rows where a country has a first author 
  
  #then, sum the total participation: for ALL authors, how many are from each country? #
  
  ungroup() %>% 
  group_by(value) %>% 
  mutate(total_percountry = case_when( row_number() == 1 ~ sum(total_percountry_peryear,na.rm=TRUE),
    .default=NA ), #get the total number of times a country is in a published paper  
    
  #then, get the total number of first author papers for each country? #
    
         total_firstauth = case_when(row_number() == 1 ~ sum(firstauth,na.rm=TRUE),
    .default=NA )
    ) %>% #get the total number of first author papers for each  
  ungroup() %>%
  select(-name) %>% distinct() %>%
  arrange(desc(total_percountry)) %>% 
  mutate(value=factor(value,levels=rev(unique(value)),ordered=TRUE)) #order in descending order by the total number of papers
```
```{r}
#check accuracy 
sum(year.countries.df$firstauth,na.rm=TRUE)
sum(year.countries.df$total_firstauth,na.rm=TRUE)
```
*How many papers did we start with/end with?*  
```{r}
nrow(spdsht)
nrow(rel.spdsht)
```
454 papers -> 317 papers  

_Regional focus_  
```{r ecodf}
eco.df <- author.countries.df %>% 
  separate_wider_delim(.,cols=countries_all, #separate the list of all author countries 
                       delim=",", #separate by comma
                       names_sep="_", #delimiter separating names of new columns
                       too_few="align_start")  %>% 
  
  rename(.,"FocusGeography"="Paper Geographical Area of Focus",
         "Management"="Marine Management  - mention/focus",
         "MPA"="Protected areas (MPAs etc.) - mention/focus",
         "MP"="Protected Area Management Plans - mention/focus" ,
         "Policy"="Policy/Law/Legislation/Framework or governance - mention/focus",
         "GeoFocus"="Focus on geomorphology/physical characteristics") %>% 
  mutate(FocusGeography=case_when(str_detect(FocusGeography,"whole region") ~ "MAR",
                                  str_detect(FocusGeography,"Wider Caribbean") ~ "Caribbean",
                                  .default=FocusGeography)) %>% 
  separate_wider_delim(.,cols="FocusGeography", #separate the list of all author countries 
                       delim=",", #separate by comma
                       names_sep="_", #delimiter separating names of new columns
                       too_few="align_start")

```

_Ecosystem focus_  
```{r geodf}
geo.df <- eco.df %>% 
  select(c(year,starts_with("countr"), #keep all the country affiliations of the authors  
           #starts_with("Focus"),
           GeoFocus,`Focus on Mangrove`,`Focus on Coral`,`Focus on Seagrass`,`Focus on  fish/fisheries`) #keep the ecosystem topics 
         ) %>% 
  pivot_longer(cols=c(GeoFocus,
                      `Focus on Mangrove`,`Focus on Coral`,
                      `Focus on Seagrass`,`Focus on  fish/fisheries`),
               values_drop_na = TRUE
               ) %>% #collapse all of the ecosystem focus options into one column  
  
  mutate(value=case_when(str_detect(value,"Unsure") ~ NA,
                         str_detect(value,"No") ~ NA,
                         str_detect(value,"N/A") ~ NA,
                         str_detect(value,"ecommendati") ~ NA, #may need to change this depending on whether we want to consider this a yes or no
                         str_detect(value,"focus") ~ "Yes",
                         str_detect(value,"only") ~ NA,
                         str_detect(value,"fish") ~ "Yes",
                         str_detect(value,"ention") ~ "Yes",
                         .default=value),
         
         name=case_when(name=="GeoFocus" ~ "Geophysical",
                        name=="Focus on Seagrass" ~ "Seagrass",
                        name=="Focus on Coral" ~ "Coral",
                        name=="Focus on Mangrove" ~ "Mangrove",
                        name=="Focus on  fish/fisheries" ~ "Fisheries",
                        .default=name))  %>% 
    
  rename("Focus"="name", 
         "FocusY"="value") %>%
  

  pivot_longer(cols=c(starts_with("countr")),
               values_drop_na = TRUE) %>% #collapse all of the countries into one column  
  #filter(!is.na(value) & !is.na(FocusY)) %>% 
  mutate(value=str_trim(value,side="both")) %>% 
  filter(!is.na(FocusY)) %>%
   group_by(year,value,Focus) %>% 
mutate(focus_peryear=n())  %>% 
  #group_by(year,value,name) %>% 
  #ungroup() %>%
  #group_by(year,value,Focus) %>% 
  #mutate(focus_peryear = n()) %>%
  select(-c(name,FocusY)) %>%
  distinct() %>% 
  ungroup() %>% group_by(value) %>% 
  mutate(n=n()) %>% #sum the total number of times a country appears in the list
  #filter(n>1) %>%
  arrange(n) %>% 
  mutate(value=factor(value,levels=unique(.$value),ordered=TRUE)) 
```
```{r geodf2}
geo.df2 <- eco.df %>% 
  select(c(`Paper #`,year,#starts_with("countr"), #keep all the country affiliations of the authors  
           starts_with("MP"),
           Policy,Management,
           GeoFocus,`Focus on Mangrove`,`Focus on Coral`,`Focus on Seagrass`,`Focus on  fish/fisheries`), 
           starts_with("Focus"),
         "Adaptive Management","Seascape","Ridge to Reef",starts_with("Resilience")
         ) %>%  pivot_longer(cols=c(starts_with("FocusGeo")), #give each geographic focus its own row (so papers occurring in multiple countries will have multiple lines) 
               values_drop_na = TRUE) %>% 
  rename("GeographicFocus"=value) %>% 
  select(-name) %>% #remove column of old column names for "focusgeography"

  mutate(GeographicFocus=str_trim(GeographicFocus,side="both")) %>% #remove leading and trailing whitespace
  ungroup() %>% 
  mutate(across(where(is.character), ~ case_when(str_detect(.,"No") ~ NA,
                         str_detect(.,"N/A") ~ NA,
                         str_detect(.,"ecommendati") ~ NA, #may need to change this depending on whether we want to consider this a yes or no
                         str_detect(.,"focus") ~ "Yes",
                         str_detect(.,"only") ~ NA,
                         str_detect(.,"fish") ~ "Yes",
                         str_detect(.,"ention") ~ "Yes",
                         .default=.))) %>% 
  select(-`Paper #`) %>%
  group_by(GeographicFocus) %>% 
mutate(nval=n())  %>% 
  group_by(GeographicFocus,year) %>%
  summarize_at(vars(-group_cols()),~sum(!is.na(.))) %>%
  ungroup() %>% rowwise() %>% 
  mutate("Alternative Livelihoods" = c_across(contains("tourism")) %>% sum(.),
         "Community Engagement" = c_across(any_of(c(contains("ngagement"),
                                                    contains("takeholder")))) %>% sum(.)
         ) %>% 
  select(.,-c("Focus on Community Engagement",contains("takeholder"),contains("tourism"))) %>%
  
 # mutate(across(-c(GeographicFocus,year), ~./nval )) %>%
select(-c(nval) ) %>% 
  pivot_longer(cols=-c(GeographicFocus,year)) %>%
  rename("Topic_all"="name",
         "Topic_all_num"="value") %>% 
  ungroup() %>% 
  
  mutate("Topic_all" = gsub("Focus on","",Topic_all) %>% str_to_title() %>% str_trim(.,side="both"),
         "Topic_all" = gsub("Mpa","MPA",Topic_all),
         "Topic_all" = gsub("Mp","ManagementPlan",Topic_all),
         
         "Topic_all" = case_when(Topic_all=="AdaptiveManagement" ~ "Adaptive Management",
                                 str_detect(Topic_all,"ender") ~ "Gender Equity",
                                 str_detect(Topic_all,"iveli") ~ "Alternative Livelihoods",
                                 str_detect(Topic_all,"nowledge") ~ "Traditional Knowledge",
                                 str_detect(Topic_all,"esilience") ~ "Resilience",
                                 
                                 str_detect(Topic_all,"Geofocus") ~ "Geophysical",
                                 str_detect(Topic_all,"oral")  ~ "Coral Reefs",
                                 str_detect(Topic_all,"ish")  ~ "Fish and Fisheries",
                                 str_detect(Topic_all,"Other Ecosystem") ~ "Other Ecosystems",
                                 
         .default=Topic_all) 
         ) %>% 
  mutate(Topic_categories = case_when(Topic_all=="Resilience" ~ "Management type",
                                      Topic_all=="Adaptive Management" ~ "Management type",
                                      Topic_all=="Seascape" ~ "Management type",
                                      Topic_all=="Ridge To Reef" ~ "Management type",
                                      
                                      Topic_all=="Mangrove" ~ "Ecosystem",
                                      Topic_all=="Coral Reefs" ~ "Ecosystem",
                                      Topic_all=="Seagrass" ~ "Ecosystem",
                                      Topic_all=="Other Ecosystems" ~ "Ecosystem",
                                      Topic_all=="Geophysical" ~ "Ecosystem",
                                      Topic_all=="Fish and Fisheries" ~ "Fish and Fisheries",
                                      
                                      str_detect(Topic_all,"ender") ~ "Community-based",
                                      str_detect(Topic_all,"ngage") ~ "Community-based",
                                      str_detect(Topic_all,"nowledge") ~ "Community-based",
                                      str_detect(Topic_all,"oluntourism") ~ "Community-based",
                                      str_detect(Topic_all,"iveli") ~ "Community-based",
                                      str_detect(Topic_all,"takeholder") ~ "Community-based",
                                      
                                      str_detect(Topic_all,"MP") ~ "Policy",
                                      str_detect(Topic_all,"Policy") ~ "Policy",
                                      str_detect(Topic_all,"Management") ~ "Policy",
                                    
                                      .default="Major topics")) %>% 
  mutate(GeographicFocus=factor(GeographicFocus,
                                levels=c("Caribbean","MAR","Mexico","Belize","Guatemala","Honduras"),
                                ordered=TRUE))
```
```{r geodfcountry}
geo.df.countryFocus <-  
  eco.df %>% 
  select(c(`Paper #`,year,starts_with("countr"), #keep all the country affiliations of the authors  
           #starts_with("Focus"),
          # GeoFocus,`Focus on Mangrove`,`Focus on Coral`,`Focus on Seagrass`,`Focus on  fish/fisheries`) #keep the ecosystem topics 
         #) %>% 
   starts_with("MP"),
           Policy,Management,
           GeoFocus,`Focus on Mangrove`,`Focus on Coral`,`Focus on Seagrass`,`Focus on  fish/fisheries`), 
           starts_with("Focus"),
         "Adaptive Management","Seascape","Ridge to Reef",starts_with("Resilience")
         ) %>% 
  select(-c(starts_with("FocusGeo"))) %>%
  
    mutate(across(where(is.character), ~ case_when(str_detect(.,"No") ~ NA,
                         str_detect(.,"N/A") ~ NA,
                         str_detect(.,"ecommendati") ~ NA, #may need to change this depending on whether we want to consider this a yes or no
                         str_detect(.,"focus") ~ "Yes",
                         str_detect(.,"only") ~ NA,
                         str_detect(.,"fish") ~ "Yes",
                         str_detect(.,"ention") ~ "Yes",
                         .default=.))) %>%
  pivot_longer(cols=c(starts_with("countr")),
               values_drop_na = TRUE) %>% #collapse all of the countries into one column  
  #filter(!is.na(value) & !is.na(FocusY)) %>% 
  mutate(Country=str_trim(value,side="both")) %>%
  select(-c(value,name,`Paper #`)) %>%
  group_by(Country,year) %>%
  summarize_at(vars(-group_cols()),~sum(!is.na(.))) %>%
  ungroup() %>% rowwise() %>% 
  mutate("Alternative Livelihoods" = c_across(contains("tourism")) %>% sum(.),
         "Community Engagement" = c_across(any_of(c(contains("ngagement"),
                                                    contains("takeholder")))) %>% sum(.)
         ) %>% 
  select(.,-c("Focus on Community Engagement",contains("takeholder"),contains("tourism"))) %>%
  group_by(Country,year) %>% 
  summarise(across(where(is.integer), ~ sum(.x, na.rm = TRUE))) %>%  
  #summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_longer(cols=-c(Country,year)) %>%
  rename("Topic_all"="name",
         "Topic_all_num"="value")  %>% 
  #ungroup() %>% 
  mutate("Topic_all" = gsub("Focus on","",Topic_all) %>% str_to_title() %>% str_trim(.,side="both"),
         "Topic_all" = gsub("Mpa","MPA",Topic_all),
         "Topic_all" = gsub("Mp","ManagementPlan",Topic_all),
         
         "Topic_all" = case_when(Topic_all=="AdaptiveManagement" ~ "Adaptive Management",
                                 str_detect(Topic_all,"ender") ~ "Gender Equity",
                                 str_detect(Topic_all,"iveli") ~ "Alternative Livelihoods",
                                 str_detect(Topic_all,"nowledge") ~ "Traditional Knowledge",
                                 str_detect(Topic_all,"esilience") ~ "Resilience",
                                 
                                 str_detect(Topic_all,"Geofocus") ~ "Geophysical",
                                 str_detect(Topic_all,"oral")  ~ "Coral Reefs",
                                 str_detect(Topic_all,"ish")  ~ "Fish and Fisheries",
                                 str_detect(Topic_all,"Other Ecosystem") ~ "Other Ecosystems",
                                 
         .default=Topic_all) 
         ) %>% 
  mutate(Topic_categories = case_when(Topic_all=="Resilience" ~ "Management type",
                                      Topic_all=="Adaptive Management" ~ "Management type",
                                      Topic_all=="Seascape" ~ "Management type",
                                      Topic_all=="Ridge To Reef" ~ "Management type",
                                      
                                      str_detect(Topic_all,"ender") ~ "Community-based",
                                      str_detect(Topic_all,"ngage") ~ "Community-based",
                                      str_detect(Topic_all,"nowledge") ~ "Community-based",
                                      str_detect(Topic_all,"oluntourism") ~ "Community-based",
                                      str_detect(Topic_all,"iveli") ~ "Community-based",
                                      str_detect(Topic_all,"takeholder") ~ "Community-based",
                                      
                                      str_detect(Topic_all,"MP") ~ "Policy",
                                      str_detect(Topic_all,"Policy") ~ "Policy",
                                      str_detect(Topic_all,"Management") ~ "Policy",
                                    
                                      .default=Topic_all)) %>% 
  mutate(
    Topic_categories=factor(Topic_categories,
                                 levels=c("Community-based","Management type","Policy","Geophysical","Fish and Fisheries","Mangrove","Coral Reefs","Seagrass","Other Ecosystems"))
    #Topic_categories=factor(Topic_categories,
    #                             levels=c("Management type","Community-based","Policy","Fish and Fisheries","Mangrove","Coral Reefs","Seagrass","Other Ecosystems","Geophysical"))
         )
```

```{r geodfregion}
geo.df.regionFocus <- eco.df %>% 
  select(c(`Paper #`,year,#starts_with("countr"), #keep all the country affiliations of the authors  
           starts_with("MP"), #MPA and MP
           Policy,Management,
           GeoFocus,`Focus on Mangrove`,`Focus on Coral`,`Focus on Seagrass`,`Focus on  fish/fisheries`, 
           starts_with("Focus"),
         starts_with("Charismatic"),
         "Major focus?", #about threats
         "Adaptive Management","Seascape","Ridge to Reef",starts_with("Resilience"))
         ) %>%  pivot_longer(cols=c(starts_with("FocusGeo")), #give each geographic focus its own row (so papers occurring in multiple countries will have multiple lines) 
               values_drop_na = TRUE) %>% 
  rename("GeographicFocus"=value) %>% 
  select(-name) %>% #remove column of old column names for "focusgeography"

  mutate(GeographicFocus=str_trim(GeographicFocus,side="both")) %>% #remove leading and trailing whitespace
  ungroup() %>% 
  mutate(across(where(is.character), ~ case_when(str_detect(.,"No") ~ NA,
                         str_detect(.,"N/A") ~ NA,
                         str_detect(.,"ecommendati") ~ "Yes", #based on notes from meeting
                         str_detect(.,"focus") ~ "Yes",
                         str_detect(.,"only") ~ NA,
                         str_detect(.,"fish") ~ "Yes",
                         str_detect(.,"ention") ~ "Yes",
                         .default=.))) %>% 
  select(-`Paper #`) %>%
  rename("Threats"="Major focus?") %>%
  group_by(GeographicFocus) %>% 
mutate(nval=n())  %>% 
  group_by(GeographicFocus,year) %>%
  summarize_at(vars(-group_cols()),~sum(!is.na(.))) %>%
  ungroup() %>% rowwise() %>% 
  mutate("Alternative Livelihoods" = c_across(contains("tourism")) %>% sum(.),
         "Community Engagement" = c_across(any_of(c(contains("ngagement"),
                                                    contains("takeholder")))) %>% sum(.),
         "Integrated Management" = c_across(any_of(c(contains("Seascape"),
                                                   contains("Ridge")))) %>% sum(.)
         ) %>% 
  select(.,-c("Focus on Community Engagement","Seascape",contains("Ridge"),
              contains("takeholder"),contains("tourism"))) %>%
  
 # mutate(across(-c(GeographicFocus,year), ~./nval )) %>%
select(-c(nval) ) %>% 
  pivot_longer(cols=-c(GeographicFocus,year)) %>%
  rename("Topic_all"="name",
         "Topic_all_num"="value") %>% 
  ungroup() %>% 
  
  mutate("Topic_all" = gsub("Focus on","",Topic_all) %>% str_to_title() %>% str_trim(.,side="both"),
         "Topic_all" = gsub("Mpa","MPA",Topic_all),
         "Topic_all" = gsub("Mp","ManagementPlan",Topic_all),
         
         "Topic_all" = case_when(Topic_all=="AdaptiveManagement" ~ "Adaptive Management",
                                 str_detect(Topic_all,"ender") ~ "Gender",#"Gender Equity",
                                 str_detect(Topic_all,"iveli") ~ "Alternative Livelihoods",
                                 str_detect(Topic_all,"nowledge") ~ "Local Knowledge", #"Traditional Knowledge",
                                 str_detect(Topic_all,"esilience") ~ "Resilience and Adaptation",
                                 
                                 str_detect(Topic_all,"MPA") ~ "Protected Areas",
                                 
                                 str_detect(Topic_all,"Geofocus") ~ "Geophysical",
                                 str_detect(Topic_all,"oral")  ~ "Coral Reefs",
                                 str_detect(Topic_all,"ish")  ~ "Fish and Fisheries",
                                 str_detect(Topic_all,"harismat")  ~ "Charismatic Species",
                                 str_detect(Topic_all,"Other Ecosystem") ~ "Other Ecosystems",
                                 
         .default=Topic_all) 
         ) %>% 
  mutate(Topic_categories = case_when(str_detect(Topic_all,"Resilience") ~ "Management type",
                                      Topic_all=="Adaptive Management" ~ "Management type",
                                      Topic_all=="Integrated Management" ~ "Management type",
                                      
                                      str_detect(Topic_all,"ender") ~ "Community-based",
                                      str_detect(Topic_all,"ngage") ~ "Community-based",
                                      str_detect(Topic_all,"nowledge") ~ "Community-based",
                                      str_detect(Topic_all,"oluntourism") ~ "Community-based",
                                      str_detect(Topic_all,"iveli") ~ "Community-based",
                                      str_detect(Topic_all,"takeholder") ~ "Community-based",
                                      
                                      str_detect(Topic_all,"MP") ~ "Policy",
                                      str_detect(Topic_all,"Protected") ~ "Policy",
                                      str_detect(Topic_all,"Policy") ~ "Policy",
                                      str_detect(Topic_all,"Management") ~ "Policy",
                                      str_detect(Topic_all,"Threats") ~ "Policy",
                                    
                                      .default=Topic_all)) %>% 
  mutate(GeographicFocus=factor(GeographicFocus,
                                levels=c("Caribbean","MAR","Mexico","Belize","Guatemala","Honduras"),
                                ordered=TRUE),
         Topic_categories=factor(Topic_categories,
                                 levels=c("Community-based","Management type","Policy","Geophysical","Fish and Fisheries","Charismatic Species","Mangrove","Coral Reefs","Seagrass","Other Ecosystems"))
         )

```
```{r}
hm.subset <- geo.df.regionFocus %>% 
  filter( Topic_all %in% c("Mangrove", "Seagrass", "Coral Reefs",
                                  "Fish and Fisheries", "Charismatic Species",
                                  "Threats",
                                  
                                  "Adaptive Management",
                                  "Integrated Management", 
                                  "Resilience and Adaptation",
                                  "Community Engagement", "Local Knowledge", "Gender",
                                  
                                  "Management", "Policy", "Protected Areas"
                                  )) %>% 
  mutate(Topic_categories=case_when(Topic_categories=="Mangrove" ~ "Ecosystems",
                                    Topic_categories=="Seagrass" ~ "Ecosystems",
                                    Topic_categories=="Coral Reefs" ~ "Ecosystems",
                                    Topic_categories=="Fish and Fisheries" ~ "Species",
                                    Topic_categories=="Charismatic Species" ~ "Species",
                                    Topic_all=="Threats" ~ "Threats",
                                    Topic_categories=="Management type" & Topic_all !="Threats" ~ "Management Approaches",
                                    Topic_categories=="Policy" & Topic_all !="Threats" ~ "Management Approaches",
                                    Topic_categories=="Community-based" ~ "Engagement",
                                    .default=Topic_categories) ) %>% 
  ungroup() %>% mutate(Topic_categories=factor(Topic_categories,levels=c("Ecosystems","Species",
                                                                         "Threats","Management Approaches",
                                                                         "Engagement"),ordered=TRUE)) %>%
  arrange(Topic_categories,Topic_all) %>% 
  mutate(Topic_all=factor(Topic_all,levels=(unique(Topic_all)%>%rev()),ordered=TRUE))
```

```{r}
#nrow(marmp)
#nrow(marmp %>% filter(Earliest_or_Latest_ManagementPlan!="Earliest" & !is.na(Earliest_or_Latest_ManagementPlan) & Earliest_or_Latest_ManagementPlan != "Not Available"))
mp.df <- marmp %>% 
  filter(Earliest_or_Latest_ManagementPlan!="Earliest" & !is.na(Earliest_or_Latest_ManagementPlan) & Earliest_or_Latest_ManagementPlan != "Not Available") %>%
  select(c(Country,
           Year_ManagementPlan,
           #starts_with("Mention"),
           Mention_AdaptiveManagement,Mention_Seascape,Mention_Resilience,
           starts_with("Action"))) %>% 
  group_by(Country) %>% mutate(TotalMPavail=n()) %>% ungroup() %>%
  group_by(Country,Year_ManagementPlan) %>% 
mutate(nval=n()) %>% #mp per year per country
  group_by(Country,nval,Year_ManagementPlan,TotalMPavail) %>% 
  summarize_at(vars(-group_cols()),~sum(!is.na(.) & . != "No" & . != "NA")) %>% 
   ungroup() %>% rowwise() %>% 
  mutate("Integrated Management" = c_across(any_of(c(contains("Seascape"),
                                                   contains("Ridge")))) %>% sum(.)
         ) %>% 
  select(-c(contains("Seascape"),contains("Ridge"),
            contains("ivelihood"),
            #contains("Scientist"),
            contains("Scale")
            )) %>%
  pivot_longer(cols=-c(nval,Country,Year_ManagementPlan,TotalMPavail)) %>%
  
  mutate(name=gsub("Mention_","",name),
         name=gsub("Mentioned_","",name),
         name=gsub("Action_","",name)) %>% 
  
  rename("Topic_all"="name",
         "Topic_all_num"="value") %>% 
  mutate("Topic_all" = case_when(str_detect(Topic_all,"Geofocus") ~ "Geophysical",
                                 str_detect(Topic_all,"oral")  ~ "Coral Reefs",
                                 str_detect(Topic_all,"Other Ecosystem") ~ "Other Ecosystems",
                                 #Mangroves: no change
                                 #Seagrass: no change
                                 
                                 str_detect(Topic_all,"ish")  ~ "Fish and Fisheries",
                                 str_detect(Topic_all,"harismat")  ~ "Charismatic Species",
                                 
                                 str_detect(Topic_all,"hreat") ~ "Threats",
                                 
                                 Topic_all=="AdaptiveManagement" ~ "Adaptive Management",
                                 #"Integrated Management"
                                 str_detect(Topic_all,"esilience") ~ "Resilience and Adaptation",
                                 #lit only: Management
                                 #lit only: Policy
                                 #lit only: Protected Areas
                              
                                 Topic_all=="CommunityEngagement" ~ "Community Engagement",
                                 str_detect(Topic_all,"cientist") ~ "Scientist Engagement",
                                 str_detect(Topic_all,"ender") ~ "Gender",
                                # str_detect(Topic_all,"iveli") ~ "Alternative Livelihoods",
                                 str_detect(Topic_all,"nowledge") ~ "Local Knowledge",
         .default=Topic_all) 
         ) %>% 
   rename("year"="Year_ManagementPlan") %>%
  mutate(Topic_categories=case_when(Topic_all=="Mangrove" ~ "Ecosystems",
                                    Topic_all=="Seagrass" ~ "Ecosystems",
                                    Topic_all=="Coral Reefs" ~ "Ecosystems",
                                    Topic_all=="Geophysical" ~ "Ecosystems",
                                    Topic_all=="Other Ecosystems" ~ "Ecosystems",
                                    
                                    Topic_all=="Fish and Fisheries" ~ "Species",
                                    Topic_all=="Charismatic Species" ~ "Species",
                                    
                                    Topic_all=="Threats" ~ "Threats",
                                    
                                    Topic_all=="Adaptive Management" ~ "Management Approaches",
                                    Topic_all=="Integrated Management" ~ "Management Approaches",
                                    str_detect(Topic_all,"esilience") ~ "Management Approaches",
                                    
                                    str_detect(Topic_all,"ngagement") ~ "Engagement",
                                    Topic_all=="Gender" ~ "Engagement",
                                    #Topic_all=="Alternative Livelihoods" ~ "Engagement",
                                    Topic_all=="Local Knowledge" ~ "Engagement",
                                    
                                    .default=Topic_all) 
         ) %>% 
  ungroup() %>% mutate(year=as.integer(year),
                       Country=factor(Country,
                                levels=c("Mexico","Belize","Guatemala","Honduras"),
                                ordered=TRUE),
                       Topic_categories=factor(Topic_categories,levels=c("Ecosystems","Species",
                                                                         "Threats","Management Approaches",
                                                                         "Engagement"),ordered=TRUE)) %>% ungroup() %>%
  arrange(Topic_categories,Topic_all) %>% 
  mutate(Topic_all=factor(Topic_all,
                          levels=(unique(Topic_all) %>% rev()),
                          ordered=TRUE)
         ) 
```

# Analysis  

### Analysis: Authorship  

##### Number of countries  
Where are most authors coming from?  

*How many countries have been involved in scientific articles regarding the MAR?*    
```{r}
length(all.countries[!is.na(all.countries)])
```
43 countries  

*Most productive countries overall? (including first and coauthors)*  
```{r topcountries}
top10 <- year.countries.df %>% arrange(desc(total_percountry)) %>% .$value %>% unique() %>% .[1:10]

top15 <- year.countries.df %>% filter(value != "NA") %>% arrange(desc(total_percountry)) %>% .$value %>% unique() %>% .[1:15]

top10
top15
```
Mexico comes in first, followed by US and UK  

Do those super productive countries change if we only look at the last 10 years? How about 5? 
```{r topcountriesyear}
top10.2015 <- year.countries.df %>% filter(year>2015) %>% group_by(value) %>% mutate(newtotal=sum(total_percountry_peryear,na.rm = TRUE)) %>% arrange(desc(newtotal)) %>% .$value %>% unique() %>% .[1:10]

top10.2020 <- year.countries.df %>% filter(year>2020) %>% group_by(value) %>% mutate(newtotal=sum(total_percountry_peryear,na.rm = TRUE)) %>% arrange(desc(newtotal)) %>% .$value %>% unique() %>% .[1:10]

top10.2015
top10.2020
```
Minor changes after US and Mexico, but nothing substantial I think.  

How productive are these countries?  
```{r}
#number of papers published by countries in the top 15 most productive category
year.countries.df %>% filter(value %in%top15) %>% 
  select(c(value,total_percountry,total_firstauth)) %>% 
  filter(!is.na(total_percountry) & !is.na(total_firstauth)) %>%
  .$total_firstauth %>% sum()

#number of papers published by countries not in that category  
year.countries.df %>% filter(!value %in%top15) %>% 
  select(c(value,total_percountry,total_firstauth)) %>% 
  filter(!is.na(total_percountry) & !is.na(total_firstauth)) %>% .$total_firstauth %>% sum()

#table for top 15 
year.countries.df %>% filter(value %in%top15) %>%
  select(c(value,total_percountry,total_firstauth)) %>% 
  filter(!is.na(total_percountry) & !is.na(total_firstauth))
```
the top 15 countries encompassed 298 of the papers, or 94% of the total papers

*Who was publishing early?*  
```{r}
year.countries.df %>% filter(year < 2006) %>% .$value %>% unique() %>% length()
year.countries.df %>% filter(year < 2006) %>% select(c(year,value,firstauth,total_percountry_peryear,total_firstauth)) %>% distinct() %>% arrange(year)
```

```{r}
year.countries.df %>% filter(Oldyear < 2006) %>% select(c(Oldyear,value,firstauth,total_percountry_peryear,total_firstauth)) %>% 
  group_by(value) %>% 
  summarize(first=sum(firstauth,na.rm=TRUE),
            total=sum(total_percountry_peryear,na.rm=TRUE))
```

```{r}
length(year.countries.df %>% filter(year < 2006) %>% .$value %>% unique())
year.countries.df %>% filter(year < 2006) %>% .$value %>% unique()
```

```{r}
length(year.countries.df %>% filter(year > 2006) %>% .$value %>% unique())
year.countries.df %>% filter(year > 2006) %>% .$value %>% unique()
```

##### Number of first author countries?  
```{r}
firstauthcountries.df %>% .$first_authors %>% unique() %>% length()
```
29 different countries were first authors  
and of those 29, the USA had the most number of first author publications closely followed by Mexico, with the UK in a distant 3rd 
```{r}
firstauthcountries.df %>% filter(!duplicated(first_authors)) %>% arrange(desc(TotalPapersCountry)) %>% .$first_authors
#double check that math is right: does TotalPapersCountry add up to the total number of papers? 
#firstauthcountries.df %>% filter(!duplicated(first_authors)) %>% .$TotalPapersCountry %>% sum()
```

##### Percent authorship    
```{r participationdf}
participation.df <- author.countries.df %>% 
  group_by(Oldyear,marfirst) %>% 
  summarize(n=n()) %>% 
  pivot_wider(names_from="marfirst",values_from="n") %>% 
  mutate(total=sum(`Not included`,Lead,Coauthor,na.rm=TRUE),
         perc=sum(Lead,Coauthor,na.rm = TRUE)/total,
         leadperc=sum(Lead,na.rm=TRUE)/total
         ) %>% arrange(desc(Oldyear))

participation.df
```
Prior to 2006, 1-6 papers focusing on the MAR were published each year   
```{r}
participation.df %>% filter(Oldyear<2006 & total>0) %>% .$total %>% range()
```
Prior to 2000, only 1-2 papers focusing on the MAR were published each year   
```{r}
participation.df %>% filter(Oldyear<2000 & total>0) %>% .$total %>% range()
```
12.5% of those papers included someone from within the region in the authorship list.   
```{r}
participation.df %>% filter(Oldyear<2000 & total>0) %>% .$perc %>% mean()
```
Publishing began to increase after 2006, breaking the double digits  
```{r}
participation.df %>% filter(Oldyear>2000 & Oldyear<2012)
```
```{r}
participation.df %>% filter(Oldyear>2006 & Oldyear<2014) %>% .$perc %>% mean(.,na.rm=TRUE)
participation.df %>% filter(Oldyear>2006 & Oldyear<2014) %>% .$leadperc %>% mean(.,na.rm=TRUE)
```

In the preceding decade, 42.49% of papers have included someone from within country, 30.91% of which were led by the first author.    

```{r}
participation.df %>% filter(Oldyear>2014 & Oldyear <2025) %>% 
  .$total %>% range(.,na.rm=TRUE)
participation.df %>% filter(Oldyear>2014 & Oldyear <2025) %>% 
  .$perc %>% mean(.,na.rm=TRUE)
participation.df %>% filter(Oldyear>2014 & Oldyear <2025) %>% 
  .$leadperc %>% mean(.,na.rm=TRUE)
```
This increased to 48.55% of papers including someone from within country, 42.8% of which were led by the first author within country.  

### Analysis: Percent access    

#### Percent basic vs percent applied  
```{r applieddf,warning=FALSE,message=FALSE}
applied.df <- author.countries.df %>% group_by(year,`Type of Paper`) %>% 
  summarize(n=n()) %>% 
  mutate(`Type of Paper`=case_when(is.na(`Type of Paper`) ~ "Unknown",
                                   .default=`Type of Paper`)) %>% 
  mutate(across(where(is.character), str_replace_all, replacement="_", pattern=" ")) %>% 
  pivot_wider(names_from=`Type of Paper`,values_from="n",values_fill=0) %>% 
  column_to_rownames(.,"year") %>% 
  mutate(total=rowSums(.,na.rm=TRUE),
         perc=round(Applied_science/total,digits=4)
         ) 
applied.df
```
```{r}
applied.df %>% filter(rownames(.)>2014) %>% .$perc %>% mean()
```

#### Percent open access  
```{r oadf}
oa.df <- author.countries.df %>% group_by(year, OA) %>% 
  summarize(n=n())%>% 
  pivot_wider(names_from=OA,values_from="n",values_fill=0) %>% 
  column_to_rownames(.,"year") %>% 
  mutate(total=rowSums(.,na.rm=TRUE),
         year=rownames(.)) %>% rowwise() %>% 
        mutate( perc=round(sum(Bronze,Gold,Green,Hybrid,na.rm=TRUE)/total,digits=4)
  ) %>% column_to_rownames(.,"year") 

oa.df
```
```{r}
oa.df %>% filter(rownames(.)>2014) %>% .$perc %>% mean()
```
```{r}
arrange(oa.df,desc(perc))[1:10,]
```

#### Percent English  
```{r langdf}
lang.df <- author.countries.df %>% group_by(year,language) %>% 
  summarize(n=n()) %>% 
  pivot_wider(names_from=language,values_from="n",values_fill=0) %>% 
  column_to_rownames(.,"year") %>% 
  mutate(total=rowSums(.,na.rm=TRUE),
         perc=round(Spanish/total,digits=4)
         ) 
lang.df
```
```{r}
mean(lang.df$perc)
```
Across all years with more than one published paper, only 3.42% of papers were published in Spanish  
```{r}
lang.df %>% filter(total>1) %>% .$perc %>% mean()
```
In the last decade, only 1.36% of papers were published in Spanish    
```{r}
lang.df %>% filter(rownames(.) > 2014) %>% .$perc %>% mean()
```

### Analysis: Ecosystem    

#### Percentage by ecosystem  
Percentage of papers focused on each ecosystem - is there a bias?  
```{r}
geo.df %>% group_by(Focus) %>% summarize(n=n(),
                                         perc=round(n/nrow(.),digits=4) )
```
```{r}
geo.df %>% group_by(year,Focus) %>% 
  summarize(n=n()) %>% 
  ungroup() %>% group_by(year) %>% 
  mutate(x=sum(n),
         perc=round(n/x,digits=4)) 
```

#### Percentage by country    
```{r}
pivot_longer(eco.df,cols=starts_with("FocusGeo")) %>% 
  mutate(RegionOfFocus=trimws(value,which="both"),.before="PaperType") %>%
  filter(.,!is.na(RegionOfFocus)) %>% 
  group_by(RegionOfFocus) %>% 
  reframe(NumberOfPapers=n()) %>% 
  arrange(desc(NumberOfPapers)) %>%  #aov(NumberOfPapers~RegionOfFocus,.) %>% TukeyHSD()
  mutate(perc=NumberOfPapers/sum(NumberOfPapers)) 
```

```{r}
geo.df %>% group_by(value,Focus) %>% 
  summarize(n=n()) %>% 
  filter(value%in%top15) #%>% aov(n~value,.) %>% TukeyHSD()
```


# 6. Countries  

## Figure components  
```{r countryauthorsdotplot,message=FALSE}
country.authors.dotplot <-  year.countries.df %>% group_by(year,value) %>% summarize(firstauth=sum(firstauth,na.rm=TRUE),total_percountry_peryear=sum(total_percountry_peryear,na.rm=TRUE)) %>% 
  ggplot(aes(x=year,y=value,color=firstauth)) + 
  geom_point(aes(size=total_percountry_peryear),alpha=0.95) + 
  scale_size_continuous(name="Number of \n authors",range=c(3,10),breaks=c(1,5,10,20,40,55)) +
  scale_color_gradientn(name="Number of \n first author \n publications",
                        colors=c("#d2ebf1","#219ebc","#176e83","#023047","#ffb703","#fb9d32","#fb8500")) + 
 scale_x_continuous(breaks=c(1990,2000,2010,2020),labels=c("Cumulative \n Prior to 1990",2000,2010,2020),expand=c(0,0)) + 
  theme_bw() + 
  ggpubr::labs_pubr() +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),axis.text.x=element_text(angle=-20,hjust=0),
        legend.position="bottom",
        panel.grid.major.y = element_line(color="grey80")
        ) + 
  guides(size=guide_legend(nrow=1,theme=theme(legend.key.size = unit(3, "point")))) +
  labs(title="Authors in papers published each year",
       caption="Points without a first author publication are colored in grey")

country.authors.dotplot

##ggsave("Figures/AuthorsByYear_dotplot.png",country.authors.dotplot,width=20,height=15)
```

In the previous figure we saw how the number of first author publications has changed, but what about overall?
```{r coauthorplot,warning=FALSE}
coauthor.plot <- author.countries.df %>% 
  group_by(year,marfirst) %>% reframe(n=n()) %>%
  pivot_wider(names_from=marfirst,values_from=n,values_fill=0) %>% 
  pivot_longer(cols=c("Not included","Lead","Coauthor")) %>% 
  ungroup() %>% summarize("year"=year,"marfirst"=name,n=value) %>%
ggplot(.) + 
  geom_line(aes(x=year,y=n,group=marfirst,color=marfirst),linewidth=2.5) + viridis::scale_color_viridis(name="Papers including \n authors from the region",discrete=TRUE,option="D",begin=0.95,end=0) + 
  theme_bw() +
  #scale_fill_manual(values=country.pal) + 
  ggpubr::labs_pubr()  +
  #scale_x_continuous(name=NULL,limits=c(1990,2025),expand=c(0,0)) + 
  scale_x_continuous(breaks=c(1990,2000,2010,2020),
                     labels=c("Cumulative \n Prior to 1990",2000,2010,2020),
                     expand=c(0,0)) + 
  theme(axis.text.x=element_text(angle=-20,hjust=0,size=10,family="Helvetica")) +
  scale_y_continuous(name="Number of papers", 
                     limits=c(0,14),expand=c(0,0.5),
                     breaks=c(0,2,4,6,8,10,12,14))

coauthor.plot

#ggsave("Figures/RegionalAuthorship_lineplot.png",coauthor.plot,width=7,height=5)
```

```{r}
coauthor.plot2 <- author.countries.df %>% 
  group_by(year,marfirst) %>% reframe(n=n()) %>% ungroup() %>%
  group_by(year) %>%
  mutate(total=sum(n)) %>% ungroup() %>%
  group_by(year,marfirst) %>% #mutate(perc=n/total) %>%
  pivot_wider(names_from=marfirst,values_from=n,values_fill=0) %>% 
  pivot_longer(cols=c("Not included","Lead","Coauthor")) %>% 
  ungroup() %>% summarize("year"=year,"marfirst"=name,n=value
                          ) %>% 
  group_by(year) %>% mutate(total=sum(n),perc=n/total) %>%
  
ggplot(.) + 
  geom_bar(aes(x=year,y=perc,#y=n,
               group=marfirst,
               #color=marfirst,
               fill=marfirst
               ),stat="identity",position="stack") +
  #geom_line(aes(x=year,y=n,group=marfirst,color=marfirst),linewidth=2.5) +
 # viridis::scale_color_viridis(name="Papers including \n authors from the region",discrete=TRUE,option="D",begin=0.95,end=0) + 
   viridis::scale_fill_viridis(name="Papers including \n authors from the region",discrete=TRUE,option="D",begin=0.95,end=0) + 
  theme_bw() +
  ggpubr::labs_pubr()  +
  scale_x_continuous(breaks=c(1990,2000,2010,2020),
                     labels=c("Cumulative \n Prior to 1990",2000,2010,2020),
                     expand=c(0,0)) + 
  theme(axis.text.x=element_text(angle=-20,hjust=0)) +
  scale_y_continuous(name="Number of papers", 
                     )

coauthor.plot2
#ggsave("Figures/RegionalAuthorship_lineplot2.png",coauthor.plot2,width=7,height=5)

```


```{r t15coauthorplot,warning=FALSE,message=FALSE}
subs.country.authors.dotplot <- year.countries.df %>% group_by(year,value) %>% summarize(firstauth=sum(firstauth,na.rm=TRUE),total_percountry_peryear=sum(total_percountry_peryear,na.rm=TRUE)) %>% filter(value%in%top15) %>% filter(!is.na(firstauth) | !is.na(total_percountry_peryear)) %>%
  ggplot(aes(x=year,y=value,
             color=ifelse(!is.na(firstauth),firstauth,NA))) + 
  geom_point(aes(size=ifelse(!is.na(total_percountry_peryear),total_percountry_peryear,NA)),alpha=0.95) + 
  scale_size_continuous(name="Number of \n authors",range=c(3,10),breaks=c(1,5,10,20,40,55)) +
  scale_color_gradientn(name="Number of \n first author \n publications",
                        colors=c("grey","#219ebc","#176e83","#023047","#ffb703","#fb9d32","#fb8500"
                                 ),
                         breaks=c(0,1,3,5,7,9),
                        labels=c(0,1,3,5,7,9)) + 
  #scale_x_continuous(minor_breaks = seq(from=1960,to=2025,by=5)) +
  theme_bw() +
  ggpubr::labs_pubr() +
  #scale_x_continuous(limits=c(1990,2026), expand = c(0,0)) + 
  scale_x_continuous(breaks=c(1990,2000,2010,2020),
                     labels=c("Cumulative \n Prior to 1990",2000,2010,2020),
                     expand=c(0,0)) + 
  theme(axis.text.x=element_text(angle=-20,hjust=0)) +
  theme(legend.position="top")+labs(caption=NULL)  +
  theme(text=element_text(size=20),
        family="Helvetica",
        axis.title=element_blank(),
        panel.grid.major.y = element_line(color="grey80")
        ) + 
  guides(size=guide_legend(nrow=1,theme=theme(legend.key.size = unit(3, "point")))) +
   guides(color=guide_colorbar(barwidth=16)) +
  labs(title="Authorship of papers published each year",
       caption="Points without a first author from that country are colored in grey")

subs.country.authors.dotplot

#ggsave("Figures/t15_AuthorsByYear_dotplot.png",subs.country.authors.dotplot,
       width=10,height=6,bg="white")
```

## PubFig: Country authorship over time   
Figure showing publication patterns for the 15 most productive countries.  
A. looks at how magnitude and participation in published research focused on the MAR has changed over time. The height of each country's ridge indicates the number of authors per year.   
B. looks at who is leading research in the MAR. Point size corresponds to the number of authors from that countries in published papers per year. The color of the point represents how many, if any, of those publications are first-author publications. 

Over time, authors from more countries were participating in research focused in the MAR and the number of publications per year increases as well. Authors from a total of 43 countries were involved in publishing research set in the MAR. Of those, 29 countries had at least one first author publication. Authors from Mexico and the United States were the most productive, with 107 and 108 first author publications, respectively. 

# 7. Accessibility  
Look at the type of paper (we're assuming applied science papers are more directly usable by managers) and accessibility (open access status) over time  

### Figure components  
```{r paperplots}
paper.lang <- ggplot(author.countries.df) + 
  geom_bar(aes(x=year,fill=language),color="black") + scale_x_continuous(breaks=c(1990,2000,2010,2020),
                     labels=c("Cumulative \n Prior to 1990",2000,2010,2020),
                     expand=c(0,0)) + 
  theme(axis.text.x=element_text(angle=-20,hjust=0)) +
  viridis::scale_fill_viridis(name="Published language",option="plasma",discrete=TRUE,begin=0.2,end=0.6)

paper.type <- ggplot(author.countries.df) + 
  geom_bar(aes(x=year,fill=`Type of Paper`),color="black")  + 
  scale_y_continuous(expand=c(0,0),limits=c(0,25)) + scale_x_continuous(breaks=c(1990,2000,2010,2020),
                     labels=c("Cumulative \n Prior to 1990",2000,2010,2020),
                     expand=c(0,0)) + 
  theme(axis.text.x=element_text(angle=-20,hjust=0)) +
  viridis::scale_fill_viridis(name="Type of research",option="plasma",discrete=TRUE,end=0.8) +
  guides(fill=guide_legend(nrow=2))

paper.oa <- ggplot(author.countries.df) + 
  geom_bar(aes(x=year,fill=OA),color="black") + 
  scale_y_continuous(expand=c(0,0),limits=c(0,25)) + scale_x_continuous(breaks=c(1990,2000,2010,2020),
                     labels=c("Cumulative \n Prior to 1990",2000,2010,2020),
                     expand=c(0,0)) + theme_bw() +
  theme(axis.text.x=element_text(angle=-20,hjust=0)) +
  scale_fill_manual(name="Open access status",
                    values=c("#335c97","#e09f3e","#a3b369","#9e2a2b","grey50"),
                    na.value = "grey60") + guides(fill=guide_legend(nrow=2)) + 
                                    ylab("Number of papers") + xlab("") + guides(fill=guide_legend(ncol=1))

paper.lang
paper.type 
paper.oa

#ggsave("Figures/Language_barplot.png",paper.lang,width=7,height=5)
#ggsave("Figures/PaperType_barplot.png",paper.type,width=7,height=5)
#ggsave("Figures/OA_barplot.png",paper.oa,width=7,height=5)
```

## PubFig: Accessibility  
```{r accessplot}
accessibility.plot <- wrap_plots((paper.type + 
                                    ylab("Number of papers") + xlab("")),
                                 (paper.oa + 
                                    ylab("Number of papers") + xlab(""))) + 
  plot_annotation(tag_levels="A",tag_suffix = ".") & theme_bw() + 
  ggpubr::labs_pubr()  + 
  theme(text=element_text(size=10),
        family="Helvetica",
        #axis.title=element_blank(),
        legend.position="bottom",
        panel.grid.major.y = element_line(color="grey80"),
        line = element_line(linewidth=0.5)
        )

accessibility.plot

#ggsave(filename="Figures/Access_cowplot.png",plot=accessibility.plot,height=5,width=15,bg="white")
```

# 8. Focus    

## Figure components    
Regional:  
```{r geofocusbarplot}
geo.plot <- pivot_longer(eco.df,cols=starts_with("FocusGeo")) %>% 
  mutate(RegionOfFocus=trimws(value,which="both"),.before="PaperType",
         RegionOfFocus=factor(RegionOfFocus,levels=c("Belize","Mexico","Honduras","Guatemala","MAR","Caribbean"))
  ) %>%
  filter(.,!is.na(RegionOfFocus)) %>% 
  ggplot(aes(x=RegionOfFocus,fill=RegionOfFocus)) +
  geom_bar(stat="count",position="stack") + 
  labs(x="Geographic Focus of Paper",
       y="Number of Papers") + 
  viridis::scale_fill_viridis(option="C",discrete=TRUE,end=0.9) + 
  guides(fill=guide_legend()) +
  theme_bw() +
  ggpubr::labs_pubr()  +
  scale_y_continuous(expand = c(0,2)) +
  #cowplot::theme_minimal_hgrid() +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),
        legend.position="none",
        panel.grid.major.y = element_line(color="grey80"),
        panel.grid.minor.x = element_line(linetype=2,color="lightgrey",linewidth = 0.25)
        ) + 
  labs(title="Geographical study area") 

geo.plot
```
```{r geofocustable}
geo.table <- pivot_longer(eco.df,cols=starts_with("FocusGeo")) %>% 
  mutate(RegionOfFocus=trimws(value,which="both"),.before="PaperType"
  ) %>%
  filter(.,!is.na(RegionOfFocus)) %>% 
  group_by(RegionOfFocus) %>% 
  reframe(NumberOfPapers=n()) %>% 
  mutate(RegionOfFocus=factor(RegionOfFocus,levels=c("Caribbean","MAR","Mexico","Belize","Guatemala","Honduras"))) %>% 
  arrange(RegionOfFocus) %>%
  #arrange(desc(NumberOfPapers)) %>% 
  flextable::flextable(.) %>% flextable::set_header_labels(RegionOfFocus="Geographical Focus",NumberOfPapers="Number of Papers") %>%
  #flextable::set_table_properties(layout="autofit") %>% 
  flextable::theme_zebra() %>%
  flextable::fontsize(size=10,part="all")

geo.table
```

Ecosystem:  
```{r topicfocustable,message=FALSE}
topic.table <- geo.df.countryFocus %>% ungroup() %>% 
  select(-c(year,Topic_all)) %>% 
  group_by(Country,Topic_categories) %>%
  summarize(Topic_all_num=sum(Topic_all_num)) %>%
  #summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_wider(names_from=Topic_categories,
              values_from=Topic_all_num) %>%
  filter(Country %in% top15) %>%
  #mutate(Total=c_across(where(is.numeric)) %>% sum(.)) %>%
  #bind_rows(summarise(.,across(where(is.numeric),sum),
  #                    across(where(is.character),~"Total"))) %>% 
  flextable::flextable(.) %>% 
  flextable::set_header_labels(.,
                               #"Geophysical"="Geo.",
                               "Coral"="CoralReefs") %>%
  flextable::theme_zebra() %>%
  flextable::set_table_properties(layout="autofit") %>% 
  flextable::fontsize(size=10,part="all") #%>% hline(i=14) %>% #vline(j=6) %>%
  #flextable::bold(i=15) %>% flextable::italic(j=7)

topic.table
```
```{r}
region.table <- 
  geo.df.regionFocus %>% ungroup() %>% 
  select(-c(year,Topic_all)) %>% 
  group_by(GeographicFocus,Topic_categories) %>%
  summarize(Topic_all_num=sum(Topic_all_num)) %>%
  #summarise(across(everything(), ~ sum(.x, na.rm = TRUE))) %>% 
  pivot_wider(names_from=Topic_categories,
              values_from=Topic_all_num) %>%
  #filter(GeographicFocus %in% top15) %>%
  #mutate(Total=c_across(where(is.numeric)) %>% sum(.)) %>%
  #bind_rows(summarise(.,across(where(is.numeric),sum),
  #                    across(where(is.character),~"Total"))) %>% 
  flextable::flextable(.) %>% 
  flextable::theme_zebra() %>%
  flextable::set_table_properties(layout="autofit") %>% 
  flextable::fontsize(size=10,part="all") #%>% hline(i=14) %>% #vline(j=6) %>%
  #flextable::bold(i=15) %>% flextable::italic(j=7)

region.table
```

```{r}
labs.hm <- c(1,2,3,4,5,6,7,8)
names(labs.hm) <- c("<1991","1991-1995", "1996-2000","2001-2005","2006-2010","2011-2015","2016-2020","2021-2025")

labs.hm10 <- c(1,2,3,4,5)
names(labs.hm10) <- c("<1991","1991-2000","2001-2010","2011-2020",">2021")
```

## PubFig: Heatmap MP comp  

### 5 year increment  
```{r marfocushm}  
MARcountries_hm_df <- hm.subset %>%
  mutate(yearBin=case_when((year>1990 & year<1996) ~ 2,
                           (year>=1996 & year<2001) ~ 3,
                           (year>=2001 & year<2006) ~ 4,
                           (year>=2006 & year<2011) ~ 5,
                           (year>=2011 & year<2016) ~ 6,
                           (year>=2016 & year<2021) ~ 7,
                           (year>=2021) ~ 8,
                           .default=1)
         ) %>% 
   group_by(GeographicFocus,
           Topic_all,
           #Topic_categories,
           yearBin) %>% 
  mutate(Topic_catBin=sum(Topic_all_num),
         yearBin=factor(yearBin,levels=c(1,2,3,4,5,6,7,8),ordered=TRUE)
         ) %>%
  distinct() %>%
  ungroup() %>% 
  complete(yearBin,
           Topic_all,
           #Topic_categories,
           GeographicFocus,
           fill=list(n=0,Topic_catBin=0,Topic_all_num=0)) %>%
  arrange(Topic_categories)  



MARcountries_heatmap <-  MARcountries_hm_df %>%
  ggplot(., aes(x=yearBin,y=Topic_all,fill=Topic_catBin)) + 
  geom_tile(group="Topic_categories",position="identity") +
facet_wrap(~GeographicFocus,nrow=1) +
  theme_bw() +
  scale_fill_gradient(name="Number \n of papers",low="#FFFFD9",high="#0B205E") + 
  scale_x_discrete(labels=names(labs.hm)) +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),
        axis.text.x=element_text(angle=90,hjust=0.85)
        ) #+labs(title="Change in manuscript focus over time by country") 
MARcountries_heatmap
#ggsave("Figures/MARcountries_heatmap.png",MARcountries_heatmap,width=15,height=7)
```

```{r}
MARmp_hm_df <- mp.df %>%
  mutate(yearBin=case_when((year>1990 & year<1996) ~ 2,
                           (year>=1996 & year<2001) ~ 3,
                           (year>=2001 & year<2006) ~ 4,
                           (year>=2006 & year<2011) ~ 5,
                           (year>=2011 & year<2016) ~ 6,
                           (year>=2016 & year<2021) ~ 7,
                           (year>=2021) ~ 8,
                           .default=1)
         ) %>% 
  group_by(Country,
           yearBin,
           Topic_all) %>% 
  mutate(Topic_catBin=sum(Topic_all_num),
    #Topic_catBin=sum(Topic_all_num),
         yearBin=factor(yearBin,levels=c(1,2,3,4,5,6,7,8),ordered=TRUE)
         ) %>%
  distinct() %>%
  ungroup() %>% 
  complete(yearBin,
           Topic_all,
           #Topic_categories,
           Country,
  #         fill=list(n=0,Topic_catBin=0,Topic_all_num=0)
  ) %>%
  arrange(Topic_categories) 


MARmp_heatmap <- MARmp_hm_df %>%
  #complete(yearBin,Topic_all,GeographicFocus,
  #         fill=list(n=0,Topic_allBin=0,Topic_all_num=0)) %>%
  #filter(yearBin < 8) %>% 
    #ggplot(., aes(x=yearBin,y=Topic_categories,fill=Topic_catBin)) + 
  ggplot(., aes(x=yearBin,y=Topic_all,fill=Topic_catBin)) + 
  #ggplot(., aes(x=yearBin,y=Topic_all,fill=Topic_all_num)) + 
  geom_tile(group="Topic_categories",position="identity") +
facet_wrap(~Country,nrow=1) +
  theme_bw() +
  scale_fill_gradient(name="Number of \n management plans",na.value = "grey45",low="#FFFFD9",high="#0B205E") + 
  scale_x_discrete(labels=names(labs.hm)) +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),
        axis.text.x=element_text(angle=90,hjust=0.85)
        ) #+labs(title="Change in manuscript focus over time by country") 
MARmp_heatmap
#ggsave("Figures/MARmp_heatmap.png",MARmp_heatmap,width=15,height=7)
```
```{r,warning=FALSE}
wrap_plots(MARcountries_heatmap,
           MARmp_heatmap,
           nrow=2) + 
  plot_annotation(tag_levels="A",tag_suffix = ".")

#ggsave("Figures/MARcountries_mp_heatmap.png",width=15,height=15)
```

### 10 year increment  
```{r marfocushm10}  
MARcountries_hm_df_10yr <- hm.subset %>%
  mutate(yearBin=case_when((year>1990 & year<2001) ~ 2,
                           (year>=2001 & year<2011) ~ 3,
                           (year>=2011 & year<2021) ~ 4,
                           (year>=2021) ~ 5,
                           .default=1)
         ) %>% 
   group_by(GeographicFocus,
           Topic_all,
           #Topic_categories,
           yearBin) %>% 
  mutate(Topic_catBin=sum(Topic_all_num),
         yearBin=factor(yearBin,levels=c(1,2,3,4,5),ordered=TRUE)
         ) %>%
  distinct() %>%
  ungroup() %>% 
  complete(yearBin,
           Topic_all,
           #Topic_categories,
           GeographicFocus,
           fill=list(n=0,Topic_catBin=0,Topic_all_num=0)) %>%
  arrange(Topic_categories)  

MARcountries_heatmap_10yr <- MARcountries_hm_df_10yr %>%
  ggplot(., aes(x=yearBin,y=Topic_all,fill=Topic_catBin)) + 
  geom_tile(group="Topic_categories",position="identity") +
facet_wrap(~GeographicFocus,nrow=1) +
  theme_bw() +
  scale_fill_gradient(name="Number \n of papers",low="#FFFFD9",high="#0B205E") + 
  scale_x_discrete(labels=names(labs.hm10))  + guides(fill=guide_colorbar(barheight=16,reverse=TRUE,title="")) +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),
        axis.text.x=element_text(angle=90,hjust=0.85)
        ) #+labs(title="Change in manuscript focus over time by country") 
MARcountries_heatmap_10yr
#ggsave("Figures/MARcountries_heatmap_10yr.png",MARcountries_heatmap_10yr,width=15,height=7)
```

```{r}
MARmp_hm_df_10yr <- mp.df %>%
  mutate(yearBin=case_when((year>1990 & year<2001) ~ 2,
                           (year>=2001 & year<2011) ~ 3,
                           (year>=2011 & year<2021) ~ 4,
                           (year>=2021) ~ 5,
                           .default=1)
         ) %>% 
  group_by(Country,yearBin, Topic_all) %>% 
  mutate(Topic_catBin=sum(Topic_all_num),
         yearBin=factor(yearBin,levels=c(1,2,3,4,5),ordered=TRUE)
         ) %>%
  distinct() %>% ungroup() %>% 
  complete(yearBin,
           Topic_all,
           Country,
  #         fill=list(n=0,Topic_catBin=0,Topic_all_num=0)
  ) %>%
  arrange(Topic_categories) 

MARmp_heatmap_10yr <- MARmp_hm_df_10yr %>%
  ggplot(., aes(x=yearBin,y=Topic_all,fill=Topic_catBin)) + 
  #ggplot(., aes(x=yearBin,y=Topic_all,fill=Topic_all_num)) + 
  geom_tile(group="Topic_categories",position="identity") +
facet_wrap(~Country,nrow=1) +
  theme_bw() +
  scale_fill_gradient(name="Number of \n management plans",low="#FFFFD9",high="#0B205E") + 
  scale_x_discrete(labels=names(labs.hm10)) +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),
        axis.text.x=element_text(angle=90,hjust=0.85)
        ) #+labs(title="Change in manuscript focus over time by country") 
MARmp_heatmap_10yr
#ggsave("Figures/MARmp_heatmap_10yr.png",MARmp_heatmap_10yr,width=15,height=7)
```

```{r,warning=FALSE}
wrap_plots(MARcountries_heatmap_10yr,
           MARmp_heatmap_10yr,
           nrow=2) + 
  plot_annotation(tag_levels="A",tag_suffix = ".")

##ggsave("Figures/MARcountries_mp_heatmap_10yr.png",width=15,height=15)
```

### Percentage   
```{r}
MARmp_hm_df_10yr_perc <- mp.df %>% 
  mutate(yearBin=case_when((year>1990 & year<2001) ~ 2,
                           (year>=2001 & year<2011) ~ 3,
                           (year>=2011 & year<2021) ~ 4,
                           (year>=2021) ~ 5,
                           .default=1)
         ) %>% 
  group_by(Country,yearBin,Topic_categories,Topic_all) %>%
  summarize(N=sum(nval),
            n=sum(Topic_all_num)) %>% 
  mutate(perc=n/N) %>%
  mutate(yearBin=factor(yearBin,levels=c(1,2,3,4,5),ordered=TRUE)
         ) %>%
  distinct() %>% ungroup() %>% 
  complete(yearBin,
           Topic_all,
           Country,
  #         fill=list(n=0,Topic_catBin=0,Topic_all_num=0)
  ) %>%
  arrange(Topic_categories) 

MARmp_heatmap_10yr_perc <- MARmp_hm_df_10yr_perc %>%
  ggplot(., aes(x=yearBin,y=Topic_all,fill=perc)) + 
  geom_tile(group="Topic_categories",position="identity") +
facet_wrap(~Country,nrow=1) +
  theme_bw() +
  scale_fill_gradient(name="Percent of \n available \n management plans",
                      low="#FFFFD9",high="#0B205E") + 
  scale_x_discrete(labels=names(labs.hm10)) +
  theme(text=element_text(size=10),
        family="Helvetica",
        axis.title=element_blank(),
        axis.text.x=element_text(angle=90,hjust=0.85)
        ) + guides(fill=guide_colorbar(barheight=16,reverse=TRUE,title=""))
MARmp_heatmap_10yr_perc 

#ggsave("Figures/MARmp_heatmap_10yr_perc.png",MARmp_heatmap_10yr_perc,width=15,height=7)
```
```{r}
MARmp_perc_long <- MARmp_hm_df_10yr_perc %>%
  select(-c(Topic_categories,N,n)) %>% 
  pivot_wider(names_from = c(Country,yearBin), #output columns
            names_sep="_",
            names_sort=TRUE,
              values_from = perc #cell values
              ) %>% 
  mutate(across(where(is.double),~round(.,digits=2))) %>%
  complete(Topic_all) %>% arrange(rev(Topic_all)) %>%
  select(!matches("_1")) %>% 
  flextable::flextable(.) %>% 
  flextable::set_header_labels(values=c("Topic",
                               rep(c(names(labs.hm10[-1])),4)
                               )) %>%
  add_header_row(colwidths = c(1,4,4,4,4),values=c("","Mexico","Belize","Guatemala","Honduras")) %>%
  set_caption("Percentage of most recent management plans addressing topic categories in 10-year increments.") %>%
  flextable::set_table_properties(layout="autofit") %>% 
  flextable::theme_zebra() %>%
  align(.,align="center",part="all") %>%
  flextable::fontsize(size=12,part="all") %>%
  hline(i=c(3,5,6,9)) %>%
  vline(j=c(1,5,9,13)) %>%
  bold(j=1)# %>% 
 # flextable::italic(j=7)
MARmp_perc_long
```

```{r}
MARmp_perc_wide <- MARmp_hm_df_10yr_perc %>%
  select(-c(Topic_categories,N,n)) %>% 
  pivot_wider(names_from = Topic_all, #output columns
              values_from = perc #cell values
              ) %>% 
  mutate(across(where(is.double),~round(.,digits=2))) %>%
  complete(Country) %>% relocate(yearBin,.before=Country) %>%
  group_by(yearBin,Country) %>%
  arrange(yearBin,Country) %>%
  filter(yearBin!="1") %>% 
  flextable::flextable(.) %>% 
  flextable::set_header_labels(yearBin="Year") %>%
  flextable::set_table_properties(layout="autofit") %>% 
  flextable::theme_zebra() %>%
  flextable::fontsize(size=10,part="all") %>%
  hline(i=c(4,8,12)) %>%
  vline(j=2) %>%
  bold(j=1:2)# %>% 
 # flextable::italic(j=7)

MARmp_perc_wide
```

```{r}
wrap_plots(MARcountries_heatmap_10yr,
           MARmp_heatmap_10yr_perc,
           nrow=2) + 
  plot_annotation(tag_levels="A",tag_suffix = ".")

#ggsave("Figures/MARcountries_mp_heatmap_10yr_perc.png",width=10,height=10)
```

```{r}
wrap_plots(MARcountries_heatmap_10yr,
           plot_spacer(),
           gen_grob(MARmp_perc_long, fit = "width"),
           nrow=3) + plot_layout(heights=c(5,1,7))

#ggsave("Figures/MARcountries_mp_table_10yr_perc.png",width=10,height=10)
```

# Session Info  
```{r}
save.image("~/Documents/GitHubRepos/MARLitReview/.RData")
sessionInfo()
```
